Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-04-24 PAPER.tex
Tue Apr 24 18:29:09 EDT 2018

Assume we have $N$ data points where
\begin{equation}
\Delta_i = \sigma_i + \gamma_i
\end{equation}
and
\begin{eqnarray}
\langle \sigma_i \rangle & = & 0 \\
\langle \sigma^2_i \rangle & = & s \\
 \langle \sigma \sigma^T \rangle & = & s I_{N \times N} \equiv S
\end{eqnarray}
\begin{eqnarray}
\langle \gamma_i \rangle & = & 0 \\
\langle \gamma^2_i \rangle & = & g \\
 \langle \gamma \gamma^T \rangle & = & g I_{N \times N} \equiv G
\end{eqnarray}
and in addition
\begin{equation}
\langle \sigma_i \gamma_j \rangle = 0
\end{equation}
That is, we have the sum of $N$ independent samples of two random variables and we wish to estimate the variance of one of the random variables in the sum; in this case, $s$.
If we want to estimate $s$ using only the inverse covariance weighting, we find
\begin{equation}
\label{eq:CaseOneICW}
\hat{s} = \half {\cal F}^{-1} \left( \Delta^T \invC \PDeriv{C}{s}  \invC \Delta \right)
\end{equation}
and plugging in the relevant quantities gives
\begin{eqnarray}
\hat{s} & = & \half \frac{2}{N} (s+g)^2 \left( \Delta^T \frac{1}{s+g} I \frac{1}{s+g} \Delta \right) \\
& = & \frac{\Delta^T \Delta}{N}
\end{eqnarray}
We note that
\begin{equation}
\langle \hat{s} \rangle = s + g
\end{equation}
so we could derive that the correct additive bias is $g$, but this requires a separate step, and the ICW by itself does not provide an unbiased answer.  This is because the covariance structure of the contaminant and the data are identical.  In this case, it is necessary to use additional information to estimate the additive bias and subtract it.

>>>>

We illustrate the way in which the quadratic estimator, Equation \ref{eq:QuadraticEstimator}, is intended to estimate the power spectrum by considering the simpler, but similar, case of estimating the variance of data.  We have specifically chosen a toy model where the data covariance is diagonal, as we expect the EoR signal to be.  We assume we have $N$ data points $\Delta_i$ which are the sum of a desired signal $\sigma_i$ and some contaminant $\kappa_i$
\begin{equation}
\Delta_i = \sigma_i + \gamma_i
\end{equation}
and
\begin{eqnarray}
\langle \sigma_i \rangle & = & 0 \\
\langle \sigma^2_i \rangle & = & s \\
 \langle \sigma \sigma^T \rangle & = & s I_{N \times N} \equiv S
\end{eqnarray}
\begin{eqnarray}
\langle \gamma_i \rangle & = & 0 \\
\langle \gamma^2_i \rangle & = & g \\
 \langle \gamma \gamma^T \rangle & = & g I_{N \times N} \equiv G
\end{eqnarray}
and in addition
\begin{equation}
\langle \sigma_i \gamma_j \rangle = 0
\end{equation}
That is, we have the sum of $N$ independent samples of two random variables and we wish to estimate the variance of one of the random variables in the sum; in this case, $s$.
If we want to estimate $s$ using only the inverse covariance weighting, we find
\begin{equation}
\label{eq:CaseOneICW}
\hat{s} = \half {\cal F}^{-1} \left( \Delta^T \invC \PDeriv{C}{s}  \invC \Delta \right)
\end{equation}
and plugging in the relevant quantities gives
\begin{eqnarray}
\hat{s} & = & \half \frac{2}{N} (s+g)^2 \left( \Delta^T \frac{1}{s+g} I \frac{1}{s+g} \Delta \right) \\
& = & \frac{\Delta^T \Delta}{N}
\end{eqnarray}
We note that
\begin{equation}
\langle \hat{s} \rangle = s + g
\end{equation}
so in this case, {\it where the covariance structure of the contaminant is identical to the signal}, it is only possible to achieve an unbiased estimate by using some additional information to estimate the additive bias and subtract it.
